---
title: "Analysis of the FDS legal status identification module"
date: today
format: 
  pdf:
      header-includes: |
        \usepackage{fontspec}
        \setmainfont{Lato}
        \usepackage{pdflscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
execute: 
  echo: false
  warning: false
  message: false
---

```{r packages}
library(labelled)
library(tibble)
library(tidyverse)
library(tidyr)
library(forcats)
library(readr)
library(readxl)
library(haven)
library(lubridate)
library(kableExtra)
library(fusen)
library(srvyr)
library(gtable)
library(rlang)
library(stringr)
library(purrr)
library(tibble)
library(readr)
library(unhcrthemes)
library(writexl)
library(data.tree)
library(DiagrammeR)
library(DiagrammeRsvg)
library(knitr)
```

```{r}
load_flat_functions("functions/flat_functions_datamanagement.qmd", envir = globalenv())
load_flat_functions("functions/flat_functions_id_paper.qmd", envir = globalenv())
```
```{r data_import}
roster_id <- read_rds("data/roster_id.rds")
ra_id <- read_rds("data/ra_id.rds")
```


# Introduction

We can take this from the EGRISS paper

Background on the FDS

# Analysis

Introduction to the analysis...

## Identification cascade

Presentation of the questions...

### South Sudan

... 

\blandscape

```{r roster_tree_ssd}
id_vars_ssd <- c(
  "ID_00", "ID_01a", "ID_01a", "ID_02", "ID_03", "ID_04",
  "ID_05", "ID_06", "ID_06b", "ID_09Ref"#,"popgroup"
  )

roster_tree_ssd <- id_tree(df = roster_id |> filter(country == "SSD"), pop = "Roster", vars = id_vars_ssd)

export_graph(ToDiagrammeRGraph(roster_tree_ssd), "roster_tree_ssd.pdf")
include_graphics("roster_tree_ssd.pdf")
```


```{r summary_ssd}
grouped_summary_table(roster_id |> filter(country == "SSD", ID_09Ref  == 1), ID_01b, ID_02, ID_03, ID_05, ID_06, ID_06b, ID_09Ref)
```
\elandscape

\blandscape

### Cameroon


```{r}
id_vars_cmr <-
  c("ID_00",
    "ID_01a",
    "ID_01a",
    "ID_02",
    "ID_03",
    "ID_04",
    "ID_05",
    "ID_06",
    "ID_06b",
    "ID_09Ref"#,
    #"popgroup"
  )

roster_tree_cmr <- id_tree(df = roster_id |> filter(country == "CMR"), pop = "Roster", vars = id_vars_ssd)
export_graph(ToDiagrammeRGraph(roster_tree_cmr), "roster_tree_cmr.pdf")
include_graphics("roster_tree_cmr.pdf")
```

\elandscape


### Pakistan

## Question outcomes

Analysis of the outcome of each question, description of findings of the three countries and profiling of those who fall outside of the right answer;

Q1: forced to flee and had to abandon:  

Q2: valid reason:  

Q3: crossing a border:  

Q4: applying for intl protection 

Q5: outcome of the application 

## Comparison proxy and random member responses...

Comparison of the answers of the HOH and self respondent. Need to filter out hoh (think we forgot this previously)

# Questions: 
Where does demographic analysis fit in?