---
title: "Legal status--FDS identification questions"
format: html
theme: cosmo  
date: today
editor: visual
execute: 
  echo: false
  warning: false
  message: false
---

```{r packages}
#| warning: false
#| message: false
library(labelled)
library(tibble)
library(tidyverse)
library(tidyr)
library(forcats)
library(readr)
library(readxl)
library(haven)
library(lubridate)
library(kableExtra)
library(fusen)
library(srvyr)
library(gtable)
library(rlang)
library(stringr)
library(purrr)
library(tibble)
library(readr)
library(unhcrthemes)
library(writexl)
library(pins)
```

```{r sourceFunctions}
load_flat_functions("functions/flat_functions_datamanagement.qmd", envir = globalenv())
```

```{r readData}
#SSD
ssd_hhroster <- read_rds("data/SSD/hhroster_clean.rds")
ssd_rmember <- read_rds("data/SSD/rmember_clean.rds")

#PAK
pak_hhroster <- read_rds("data/PAK/hhroster_merge.rds")

#CMR
source(".Renviron")
board <- board_connect(auth = "envvar")
cmr_df_l <- pin_read(board, "nice@unhcr.org/FDS_CMR_df_l")
```

## Analysis
### South Sudan
Combine "recognized refugee" and "refugee enrollment documents"

Overview of identification questions

National of host country

```{r}
grouped_summary_table(ssd_hhroster |> filter(ageYears >= 15, ID_00 == 1), ID_01comb, ID_02simp, ID_03simp, ID_04, ID_05, ID_06, ID_06bsimp, popgroup)
```

National of other country
```{r}
grouped_summary_table(ssd_hhroster |> filter(ageYears >= 15, ID_00 == 2), ID_00, ID_01b, ID_02simp, ID_03simp, ID_05, ID_06, ID_06bsimp, popgroup)
```
Comparing ID document and proxy classification

Final outcome

Comparing outcome of responses from household head and identification document.

Inclusion error: Classified as refugees: e.g. 310 who have no documents (we don't know if this is really an error, many may have really misplaced documents).

Exclusion error: Not classified as refugees: e.g. 1473 refugees classified as foreigners/600 refugees classified as missing/132 refugees classified as asylum-seekers
```{r warning = FALSE, message = FALSE}
ssd_hhroster_15plus <- ssd_hhroster |> filter(ageYears >= 15)

popgroup_ID_09 <- table(ssd_hhroster_15plus$popgroup |> to_factor(), 
                        ssd_hhroster_15plus$ID_09 |> to_factor())

df_popgroup_ID_09 <- data.frame(popgroup_ID_09) |> 
  mutate(Var2 = case_when(Var2 == "" ~ "Missing",
                          TRUE ~ Var2)) |> 
  pivot_wider(names_from = "Var1", values_from = "Freq") %>%
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"Total"))) %>%
  mutate(Total = rowSums(.[-1])) |> 
  rename("Identification document" = "Var2")
df_popgroup_ID_09
#write_xlsx(df_popgroup_ID_09, "Identitification questions crosstabs.xlsx")
```
Identification questions and identity document

Non-nationals

```{r}
grouped_summary_table(ssd_hhroster |> filter(ageYears >= 15), ID_00) 
```
Exclusion error due to ID_01b (forced to flee from country of origin): 1220 individuals with refugee documents and 143 individuals with refugee enrollment document who were not forced to flee.

Could be misunderstanding due to specific context in South Sudan. Question reworded for next countries.

```{r warning = FALSE, message = FALSE}
grouped_summary_table(ssd_hhroster |> filter(ageYears >= 15, ID_00 == 2), ID_00, ID_01b)
```
Exclusion error due to ID_02 (main reason for fleeing): 245 individuals with refugee documents and 28 individuals with refugee enrollment document whose reason for fleeing provided in the FDS does not classify them as refugees.

```{r warning = FALSE, message = FALSE}
grouped_summary_table(ssd_hhroster |> filter(ageYears >= 15, ID_00 == 2, ID_01b == 1), ID_00, ID_01b, ID_02simp, ID_09)
```
Exclusion error due to ID_03 (crossed an international border): 93 individuals with refugee documents and 3 individuals with refugee enrollment document who did not cross an international border after fleeing.

```{r warning = FALSE, message = FALSE}
grouped_summary_table(ssd_hhroster |> filter(ageYears >= 15, ID_00 == 2, ID_01b == 1, ID_02simp == 1), ID_00, ID_01b, ID_03simp)
```
Exclusion error due to ID_05 (applied for international protection): 430 individuals with refugee documents and 30 individuals with refugee enrollment document who did not apply for international protection, classified as missing. 135 refugees falsely classfied as AS

```{r warning = FALSE, message = FALSE}
grouped_summary_table(ssd_hhroster |> filter(ageYears >= 15, ID_00 == 2, ID_01b == 1, ID_02simp == 1, ID_03simp == 1), ID_05)
```
Exclusion error due to ID_06 (refugee application status): 7 individuals who say they are waiting for response on asylum application but have refugee enrollment document

Inclusion error as refugee: 61 individuals with refugee status granted, but document is for asylum applicant

```{r warning = FALSE, message = FALSE}
grouped_summary_table(ssd_hhroster |> filter(ageYears >= 15, ID_00 == 2, ID_01b == 1, ID_02simp == 1, ID_03simp == 1, ID_05 == 1), ID_00, ID_01b, ID_06bsimp, popgroup, ID_09)
```

```{r}
grouped_summary_table(ssd_hhroster |> filter(ageYears >= 15, !is.na(popgroup)), ID_09)
```
Individual (random adult) and proxy responses

ID_00: 33 cases where proxy identifies individual as national of other country but individual state they are South Sudanese, 6 cases of the opposite

```{r warning = FALSE, message = FALSE}
IDcrosstab <- function (var1, var2) {

  tab <- table(ssd_rmember[[var1]] |> 
          to_factor(),
        ssd_rmember[[var2]] |> 
          to_factor())  |>
    kable(format = 'html') |>
    kable_styling() |>
    scroll_box(width = "100%",  box_css = "border: 0px;")
  
  tab

}

IDcrosstab("ID_00", "ID_00_random")
```

ID_01a
```{r warning = FALSE, message = FALSE}
IDcrosstab("ID_01a", "ID_01a_random")
```

ID_01b: Fairly high dispcrepency here, matches high inclusion/exclusion error as above
```{r warning = FALSE, message = FALSE}
IDcrosstab("ID_01b", "ID_01b_random")
```

ID_02
```{r warning = FALSE, message = FALSE}
IDcrosstab("ID_02", "ID_02_random")
```

ID_03
```{r warning = FALSE, message = FALSE}
IDcrosstab("ID_03", "ID_03_random")
```

ID_05
```{r warning = FALSE, message = FALSE}
IDcrosstab("ID_05", "ID_05_random")
```

ID_06b
```{r warning = FALSE, message = FALSE}
IDcrosstab("ID_06b", "ID_06b_random")
```

Individual responses and identity document (random adult)

ID_01b
```{r warning = FALSE, message = FALSE}
grouped_summary_table(ssd_rmember |> filter(ID_00_random == 2), ID_00_random, ID_01b_random, ID_09_random)
```

Exclusion error due to ID_02 (main reason for fleeing)
```{r warning = FALSE, message = FALSE}
grouped_summary_table(ssd_rmember |> filter(ID_00_random == 2, ID_01b == 1), ID_00_random, ID_01b_random, ID_02_random, ID_09_random)
```

Exclusion error due to ID_03 (crossed an international border)
```{r warning = FALSE, message = FALSE}
grouped_summary_table(ssd_rmember |> filter(ID_00_random == 2, ID_01b_random == 1, ID_02_random == 1), ID_03_random, ID_09_random)
```

Exclusion error due to ID_05
```{r warning = FALSE, message = FALSE}
grouped_summary_table(ssd_rmember |> filter(ID_00_random == 2, ID_01b_random == 1, ID_02_random == 1, ID_03_random == 1), ID_05_random, ID_09_random)
```
Exclusion error due to ID_06 (refugee application status)
```{r warning = FALSE, message = FALSE}
grouped_summary_table(ssd_rmember |> filter(ID_00_random == 2, ID_01b_random == 1, ID_02_random == 1, ID_03_random == 1, ID_05_random == 1), ID_00_random, ID_01b_random, ID_06b_random, ID_09_random)
```

