---
title: "Bias from criteria-based refugee identification"
date: "`r format(Sys.Date(),  '%d %B %Y')`"
output:
  unhcrdown::docx_simple
    #back_html: true
paged-footnotes: true
links-to-footnotes: true

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.cap = TRUE, fig.width = 6.27)
```

```{r packages}
pacman::p_load(labelled, tibble, tidyverse, forcats, fusen, srvyr, kableExtra, rlang, stringr, purrr, unhcrthemes, writexl, knitr, scales, extrafont, scales, knitr, survey, broom, prettyunits, gtools, flextable, officedown, officer, bookdown, unhcrthemes) 

loadfonts(device = "win")
```

```{r data_import}
roster_id <- read_rds("data/roster_id.rds") |>
  filter(ageYears >= 15) |>
  filter(ID_09Ref == "1")
ra_id <- read_rds("data/ra_id.rds")  |>
  filter(ageYears >= 15)  |>
  filter(ID_09Ref == "1")
```

```{r}
# Refugee based on criteria
roster_id <- roster_id |>
  mutate(criteria_ref = case_when(ID_06b == 1 ~ "Refugee based on criteria",
                                  TRUE ~ "Non-refugee based on criteria"),
         criteria_ref_num = case_when(criteria_ref == "Refugee based on criteria" ~ 1, TRUE ~ 0))
  
ra_id <- ra_id |>
  mutate(criteria_ref = case_when(ID_06b == 1 ~ "Refugee based on criteria",
                                  TRUE ~ "Non-refugee based on criteria"),
         criteria_ref_num = case_when(criteria_ref == "Refugee based on criteria" ~ 1, TRUE ~ 0))
```

# Introduction

An accurate and standardized approach to identifying refugees, asylum-seekers, and internally displaced persons (IDPs) is critical for their inclusion in national statistical systems. The Expert Group on Refugee, Internally Displaced and Stateless Persons Statistics (EGRISS) has addressed this challenge through the International Recommendations on Refugee and IDP Statistics (IRRS/IRIS), endorsed by the UN Statistics Division in 2018-2020. Rather than relying on direct self-identification, which can lead to misreporting, the EGRISS framework employs indirect criteria-based questions covering citizenship, migration reasons, and border crossings to identify displaced populations.

UNHCR's Forced Displacement Survey (FDS) implemented these guidelines in its three pilot surveys in South Sudan, Pakistan, and Cameroon, conducted across 2023 and 2024. The findings from these surveys provide an opportunity to evaluate the criteria-based identification questions across diverse refugee contexts. In a previous paper, we show that despite holding legal refugee status, many card holders fail to respond to criteria-based questions as expected.^[link] The analysis shows a range of challenges correctly identifying refugees with the criteria-based questions, for example due misunderstandings of key terminology, complexities of displacement biographies and family reunification, in multi-generational displacement contexts, and country-specific policy environments.

The misidentification warrants further examination because it appears to be non-random. Our previous analysis shows that certain groups, for example children of refugees, households which were separated and subsequently reunited, and those in fluid border regions are systematically excluded. Relying on identification questions as implemented in the FDS pilot countries would therefore misrepresent the population share of important population groups. Table X presents a summary of key socio-demographic statistics among refugees with a valid document for the three countries, disaggregating by those who complete and those who do not complete the identification cascade.

```{r}
row_order <- c("Year of arrival (median)", "Age (median)", "Female", "Separated households", "Citizenship (roster_id countries)", "Household size (mean)", "n")

# Function to summarize data for one country
summarize_country <- function(df, country_name) {
  filtered_data <- df %>% filter(Country == country_name)

  summary_df <- filtered_data %>%
    group_by(criteria_ref) %>%
    summarise(
      `Year of arrival (median)` = median(arrivalYear, na.rm = TRUE),
      `Age (median)` = median(ageYears, na.rm = TRUE),
      `Separated households` = paste0(round(mean(`Separated households`, na.rm = TRUE) * 100, 1), "%"),
      `Female` = paste0(round(mean(Female, na.rm = TRUE) * 100, 1), "%"),
      `Household size (mean)` = round(mean(hhsize, na.rm = TRUE), 1),
      n = sum(!is.na(arrivalYear)),
      .groups = "drop"
    ) %>%
    mutate(across(everything(), ~as.character(.))) %>%
    pivot_longer(
      cols = c(`Year of arrival (median)`, `Age (median)`, `Separated households`, `Female`, `Household size (mean)`,  n),
      names_to = "Variable",
      values_to = "value"
    )

  top_citizenships <- filtered_data %>%
    filter(!is.na(citizenship)) %>%
    count(citizenship) %>%
    mutate(pct = round(n / sum(n) * 100, 1)) %>%
    arrange(desc(n)) %>%
    filter(pct > 5) %>%
    slice_head(n = 3) %>%
    mutate(label = paste0(citizenship, " (", pct, "%)")) %>%
    pull(label) %>%
    paste(collapse = ", ")

  citizenship_row <- filtered_data %>%
    distinct(criteria_ref) %>%
    mutate(
      Variable = "Citizenship (roster_id countries)",
      value = top_citizenships
    )

  bind_rows(summary_df, citizenship_row) %>%
    mutate(Variable = factor(Variable, levels = row_order)) %>%
    arrange(Variable) %>%
    pivot_wider(names_from = criteria_ref, values_from = value) %>%
    flextable() %>%
    autofit()
}

```

*[COMBINE TABLES WHEN LAYOUTING]*
```{r tab-ssd, tab.cap="Summary statistics, South Sudan"}
roster_id %>%
  summarize_country(., country_name = "South Sudan")  |>
  fontsize(size = 9, part = "all") |> 
  bold(part = "head")
```

```{r tab-cmr, tab.cap="Summary statistics, Cameroon"}
roster_id %>%
  summarize_country(., country_name = "Cameroon")  |>
  fontsize(size = 9, part = "all") |> 
  bold(part = "head")
```

```{r tab-pak, tab.cap="Summary statistics, Pakistan"}
roster_id %>%
  summarize_country(., country_name = "Pakistan")  |>
  fontsize(size = 9, part = "all") |> 
  bold(part = "head")
```

In this paper, we aim to understand the nature of the bias resulting from identification using the criteria-based approach versus document-based approach for refugee identification in household surveys. Understanding and correcting these biases is essential to validate the criteria-based approach and contribute to accurate methods of statistical inclusion of displaced populations. 

Limitation: effect of inclusion error is not possible to measure from FDS data. This would be error from identification of non-refugees based on criteria who in fact do not have actual status. These people are not interviewed in FDS, however we may be able to get some frequency estimates for this type of bias from people excluded from FDS after the ID module due to failure to provide a refugee document. 

In this paper, we use three complementary empirical strategies to assess the misidentification bias in criteria-based refugee identification.  First, we model the likelihood of cascade non-completion using demographic and displacement history variables to determine whether exclusion is systematically correlated with specific household or individual characteristics. Second, we conduct difference-in-means tests comparing key outcome indicators between samples identified through the criteria-based cascade versus those identified via documentation, and alternatively between those completing the cascade versus those exiting at various stages. Third, we implement simulation analyses that exclude different household types to estimate the average deviation of outcome and protection indicators from their true values (based on documented status), thereby quantifying the systematic nature of identification bias and its implications for population-level estimates.

In the FDS, the identification questions first appear in the roster, where the head of household responds about himself/herself and about all members of the household over 15 as a proxy. The direct respondent then receives the same set of questions. In South Sudan and Cameroon, the direct respondent was the randomly selected adult, while in Pakistan it was the randomly selected woman. If the randomly selected adult/woman is the head of household who already responded as a proxy, the questions are only asked in the roster.

# Likelihood of exclusion from the identification sequence

To better understand the relationship between the likelihood of completing the refugee identification questions as expected given legal documentation and a series of demographic and contextual variables including sex, household head status, country of birth, years since arrival, separation status, and refugee location (the sample stratum), we run a series of simple logistic regression model. The results are presented as odds ratios (OR) rather than log-odds coefficients for easier interpretation. These show the change in odds of meeting refugee criteria associated with each explanatory variable.^[An OR of 1 indicates no effect, values greater than 1 indicate increased odds (e.g., 2 means double the odds), and values less than 1 indicate decreased odds (e.g., 0.5 means half the odds).]

**Key findings**

- Separation: Positive effect on likelihood of correct identification in South Sudan, negative in Pakistan and Cameroon (surpising in SSD given descriptives that separated households more likely to state that they fled for family reasons. Interact with position in household?)

- Location/stratum is consistently the strongest predictor. In Pakistan strong negative correlation with in RV and urban. In Cameroon, positive correlation with urban areas. In Cameroon, positive correlation with peri-urban and urban areas. In South SUdan very high correlation, refugees in Maban much much more likely to be identified as refugees than those in Pariang.

- Individual demographics (sex, household head status) are generally not significant

- Years since arrival: Born in host country has strong negative correlation in Pakistan and Cameroon, as expected. In Cameroon, longer stays in country correlated with higher likelihood of completing cascade as expected, in Pakistan and South Sudan mostly insignificant.

- Relationship to household head: No siginificant relationship with likelihood of completing identification cascade.

- Education: In Pakistan, higher education of HoH correlated with lower likelihood of correct identification - likely due to younger cohorts being better educated but less likely to be born abroad. In South Sudan no significant relationship, in Cameroon higher levels of education of the HoH are positively correlated with likelihood of completing identification sequence.

```{r}
countries <- unique(roster_id$Country)
outcome_var <- "criteria_ref_num"
covariates  <- c("sex",
                 "sexHoH",
                 #"hhead",
                 "Years since arrival",
                 "separated",
                 "stratum",
                 "HH_03",
                 "eduLevelHoH")

```

```{r}
roster_id <- roster_id |> 
  filter(!cob %in% c("Don't know", "Refuse to answer", "Stateless"))

roster_id <- roster_id %>%
  mutate(years_in_hc = 2024 - as.numeric(arrivalYear)) |>
  mutate(`Years since arrival` = case_when(
    cob == "Host country" ~ "Born in host country", 
    years_in_hc < 5 ~ "<5",
    years_in_hc >= 5 & years_in_hc < 10  ~ "5-9",
    years_in_hc >= 10 & years_in_hc < 20  ~ "10-19",
    years_in_hc >= 20  ~ "20+",
    TRUE ~ NA_character_
  )) |>
  mutate(eduLevelHoH = labelled::to_factor(eduLevelHoH))

roster_id$stratum <- fct_relevel(roster_id$stratum, levels(roster_id$stratum)[3], levels(roster_id$stratum)[1])


roster_id <- roster_id %>% 
  mutate(
    # Convert to factor with household head first, group uncommon categories as Other
    HH_03 = case_when(
      HH_03 == "The household head himself/herself" ~ "Household head",
      HH_03 == "Biological child" ~ "Biological child",
      HH_03 == "Wife/Husband/Partner" ~ "Wife/Husband/Partner",
      HH_03 == "Mother/Father" ~ "Mother/Father",
      HH_03 == "Sister/Brother" ~ "Sister/Brother",
      HH_03 == "Grandchild" ~ "Grandchild",
      HH_03 %in% c("Brother/Sister-in-law", 
                   "Cousin", 
                   "Father/Mother-in-law",
                   "Grandfather/mother",
                   "Niece/Nephew",
                   "Son/Daughter-in-law",
                   "Step child/Adopted child",
                   "Uncle/Aunt") ~ "Other",
      TRUE ~ "Other"
    ),
    HH_03 = factor(HH_03, levels = c("Household head", 
                                     "Wife/Husband/Partner", 
                                     "Biological child",
                                     "Mother/Father", 
                                     "Sister/Brother", 
                                     "Grandchild", 
                                     "Other"))
  )


# Helper: backtick unsafe names (spaces, punctuation)
bt <- function(x) {
  ifelse(make.names(x) != x, paste0("`", x, "`"), x)
}

# Significance stars
add_significance <- function(p_value) {
  ifelse(p_value < 0.001, "***",
         ifelse(p_value < 0.01,  "**",
                ifelse(p_value < 0.05, "*",
                       ifelse(p_value < 0.1, ".", ""))))
}

# ---- Main per-country function ----
model_refugee_probability <- function(data,
                                    country_name,
                                    outcome = "criteria_ref_num",
                                    covars = covariates,
                                    docvar = "D",             # set to NULL if you don't have D
                                    seed = 123,
                                    return_type = c("all", "summary", "odds_ratios", "model", "tidy")) {

  set.seed(seed)
  return_type <- match.arg(return_type)

  # Filter to a single country
  country_data <- data %>% filter(.data$Country == country_name)
  if (nrow(country_data) == 0) stop(paste("Country", country_name, "not found in data"))

  # If docvar provided, restrict the modeling to documented
  if (!is.null(docvar)) {
    country_data <- country_data %>% filter(.data[[docvar]] == 1)
  }

  # Drop missing on variables we use
  need_vars <- c(outcome, covars)
  country_data <- country_data %>%
    select(any_of(c("Country", need_vars))) %>% na.omit()

  # Build formula safely (support spaces)
  fml <- as.formula(
    paste0(bt(outcome), " ~ ", paste(bt(covars), collapse = " + "))
  )

  # ---- 1) Fit logistic regression model ----
  model <- glm(fml, data = country_data, family = binomial())

  # ---- 2) Odds ratios table (flextable) ----
  or_tbl <- tidy(model, conf.int = TRUE) %>%
    select(term, estimate, std.error, conf.low, conf.high, p.value) %>%
    mutate(
      odds_ratio = exp(estimate),
      or_ci_low  = exp(conf.low),
      or_ci_high = exp(conf.high),
      odds_ratio_r = round(odds_ratio, 2),
      or_ci_low_r  = round(or_ci_low, 2),
      or_ci_high_r = round(or_ci_high, 2),
      std.error_r  = round(std.error, 2),
      p.value_r    = round(p.value, 2),
      stars        = add_significance(p.value),
      or_stars     = paste0(odds_ratio_r, stars)
    ) %>%
    select(term, or_stars, std.error_r, or_ci_low_r, or_ci_high_r, p.value_r)

  or_flex <- flextable(or_tbl) %>%
    set_caption(paste("Logistic Regression (Odds Ratios) -", country_name)) %>%
    fontsize(size = 9, part = "all") |> 
    set_header_labels(
      term = "Variable",
      or_stars = "Odds Ratio",
      std.error_r = "Std. Error",
      or_ci_low_r = "OR CI Lower",
      or_ci_high_r = "OR CI Upper",
      p.value_r = "P-value"
    ) %>%
    autofit() %>%
    add_footer_lines("*** p<0.001, ** p<0.01, * p<0.05, . p<0.1")

  # ---- 3) Model diagnostics ----
  # Predicted probabilities
  country_data$predicted_prob <- predict(model, type = "response")
  
  # Distribution of predicted probabilities
  prob_plot <- ggplot(country_data, aes(x = predicted_prob, fill = factor(.data[[outcome]]))) +
    geom_histogram(position = "identity", alpha = 0.6, bins = 30) +
    labs(x = "Predicted P(Refugee = 1)", 
         y = "n",
         fill = "", 
         title = "") +
    theme_minimal() +
    scale_fill_manual(values = c("0" = "lightblue", "1" = "darkblue"))
  
  # Model fit statistics
  model_stats <- tibble(
    country = country_name,
    n_obs = nrow(country_data),
    n_refugees = sum(country_data[[outcome]] == 1),
    refugee_rate = mean(country_data[[outcome]] == 1),
    aic = AIC(model),
    bic = BIC(model),
    deviance = deviance(model),
    null_deviance = model$null.deviance,
    pseudo_r2 = 1 - (deviance(model) / model$null.deviance)
  )

  # ---- 4) Assemble outputs ----
  result <- list(
    country = country_name,
    model = model,
    odds_ratios_flextable = or_flex,
    odds_ratios_tidy = or_tbl,
    model_stats = model_stats,
    diagnostics = list(
      predicted_probabilities = country_data$predicted_prob,
      prob_distribution_plot = prob_plot
    )
  )

}

```



\newpage

```{r}
res_south_sudan <- model_refugee_probability(
  data = roster_id,
  country_name = "South Sudan",
  outcome = outcome_var,
  covars = covariates,
  docvar = NULL,
  return_type = "odds_ratios"
)

res_south_sudan$odds_ratios_flextable
```

```{r, fig.cap= "Predicted Probabilities - South Sudan"}
res_south_sudan$diagnostics$prob_distribution_plot
```

\newpage

```{r}
res_pak <- model_refugee_probability(
  data = roster_id,
  country_name = "Pakistan",
  outcome = outcome_var,
  covars = covariates,
  docvar = NULL,
  return_type = "odds_ratios"
) 

res_pak$odds_ratios_flextable
```

```{r, fig.cap= "Predicted Probabilities - Pakistan"}
res_pak$diagnostics$prob_distribution_plot

```

\newpage

```{r}
res_cmr <- model_refugee_probability(
  data = roster_id,
  country_name = "Cameroon",
  outcome = outcome_var,
  covars = covariates,
  docvar = NULL,
  return_type = "odds_ratios"
)

res_cmr$odds_ratios_flextable
```

```{r, fig.cap= "Predicted Probabilities - Cameroon"}
res_cmr$diagnostics$prob_distribution_plot

```


[age and arrival year likely highly correlated so just chose one]





# Comparison of outcome indicators between criteria-based and document-based identification

```{r}
# Function to calculate mean values
summary_ci <- function(data, var) {
  summary_criteria <- data %>%
    group_by(Country, stratum, criteria_ref) %>%
    summarise(
      mean = mean(.data[[var]], na.rm = TRUE),
      se = sd(.data[[var]], na.rm = TRUE) / sqrt(sum(!is.na(.data[[var]]))),
      n = sum(!is.na(.data[[var]])),
      .groups = "drop"
    ) %>%
    mutate(
      lower = mean - 1.96 * se,
      upper = mean + 1.96 * se,
      variable = var
    )
  
  summary_ref <- data %>%
    group_by(Country, stratum) %>%
    summarise(
      mean = mean(.data[[var]], na.rm = TRUE),
      se = sd(.data[[var]], na.rm = TRUE) / sqrt(sum(!is.na(.data[[var]]))),
      n = sum(!is.na(.data[[var]])),
      .groups = "drop"
    ) %>%
    mutate(
      lower = mean - 1.96 * se,
      upper = mean + 1.96 * se,
      variable = var
    ) |>
    mutate(criteria_ref = "Recognised refugees")
  
  bind_rows(summary_criteria, summary_ref)
}


# Function to plot mean values
plot_fun <- function(data, theme_func = theme_unhcr, fill_scale_func = scale_fill_unhcr_d) {
  ggplot(data, aes(x = stratum, y = mean, fill = criteria_ref)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
    geom_errorbar(aes(ymin = lower, ymax = upper),
                  position = position_dodge(width = 0.9),
                  width = 0.25) +
    # geom_text(aes(label = n),
    #           position = position_dodge(width = 0.9),
    #           vjust = -0.5,
    #           size = 3) +
    scale_y_continuous(
      expand = c(0, 0),
      labels = scales::percent
    ) +
    labs(title = "", y = "", x = "", fill = "") +
    theme_func(grid = "Y") +
    fill_scale_func() +
    facet_wrap(~Country , scales = "free")
}
```




The previous Methodological Note examines for each question in the identification sequence plausible explanations and shows that dropping out of the sequence is in many cases non-random. For example, respondents living in border regions and in a camp/site/refugee village are less likely to say that they ever had to cross a border, despite living outside the country of their birth. For a systematic analysis of the demographic, socio-economic and displacement-related factors that may be correlated with completing the identification sequence as expected, we conduct a logistic regression.
 
The previous analyis also discussed that certain groups can by certain definition not be classified as refugees based on the existing criteria - for example as children of refugees were never forced to flee but are likely to have (derivative) refugee status.


**Financial services**
```{r}
ra_id <- ra_id %>%
  mutate(`Financial account` = case_when(
   FI01 == 1 | FI02 == 1 | FI10 == 1 ~ 1,
   FI01 == 2 & FI02 == 2 & FI10 %in% c(2,3) ~ 0,
    TRUE ~ NA                # For individuals under 18 and all other cases, assigns NA
  ))

ra_id <- ra_id %>%
   mutate(`Financial account` = labelled(`Financial account`,
                                labels = c(
                                  "Yes" = 1,  
                                  "No" = 0
                                ),
                                label = "Proportion of people above with an account at a bank or other financial institution or with a mobile-money-service provider"
  ))


```

**Land access**
```{r}
roster_id <- roster_id |>
  mutate(
    documents = ifelse(rowSums(across(starts_with("Land10"), ~ .x == 1), na.rm = TRUE) > 0, 1, 0),
    bequeath = ifelse(rowSums(across(starts_with("Land14"), ~ .x == 1), na.rm = TRUE) > 0, 1, 0),
    tenure = ifelse(rowSums(across(starts_with("Land15"),  ~ .x %in% c(1, 2)), na.rm = TRUE) > 0, 1, 0)
  )

roster_id <- roster_id |>
mutate(`Land documents` = case_when(documents == 1~ 1,  
    TRUE ~ 0)) |>
   mutate(`Land documents` = labelled(`Land documents`,
                                labels = c(
                                  "Has Legal document" = 1,  
                                  "No legal documents" = 0  
                                ),
                                label = "People (adults) with legally recognized documentation over land")) |>
  mutate(`Land tenure` = case_when(bequeath == 1 &
                                tenure == 1 ~ 1, 
                              TRUE ~ 0))
roster_id <- roster_id |>
   mutate(`Land tenure` = labelled(`Land tenure`,
                                labels = c(
                                  "Tenure secure" = 1,  
                                  "Tenure unsecure" = 0   
                                ),
                                label = "People (adults) who perceive their rights as secure"))

land_docs_ci <- summary_ci(roster_id, "Land documents")
land_tenure_ci <- summary_ci(roster_id, "Land tenure")
```


**Informal employment**

```{r, eval=TRUE}
#Step 1: Determine who is employed based on various conditions 
ra_id <- ra_id %>%
  mutate(employed = case_when(
    EMP01 == 1 ~ 1,  # Paid employment - employees
    (EMP02 == 1 & EMP06b == 3)  ~ 1,  # Self-employment
    (EMP03 == 1 & EMP06b == 3)  ~ 1,  # Unpaid contributing family workers
    EMP04 == 1  ~ 1,  # Absent but still employed
    (EMP05 == 1 & EMP06a == 3)  ~ 1,  # Absent but self-employed
    (EMP02 == 1 & EMP06b %in% c(1, 2) & EMP06c %in% c(1, 2))  ~ 1,  # Farming/rearing/fishing for sale
    (EMP05 == 1 & EMP06a %in% c(1, 2) & EMP06c %in% c(1, 2))  ~ 1,  # Absent but farming for sale
    TRUE ~ 0
  ))

#Step 2: Status in employment based on isce93 
ra_id <- ra_id %>%
  mutate(isce93 = case_when(
    employed== 1 & EMP09 %in% c(1,2,3) ~ 1, #Employees
    employed== 1 & EMP09 == 4 & EMP10_7 > 1 ~ 2, #Employers - assuming that all workers are paid
    employed== 1 & EMP09 == 4 & EMP10_7 == 1 ~ 3, #Own-account workers
    employed== 1 & EMP09 == 5 ~ 4, # Contributing family workers
    employed== 1 & (EMP09 == 96 | is.na(EMP09)) ~ 5, #Workers not classifiable by status
    employed == 0 ~ NA_real_,  # Set to NA when employed == 0
    TRUE ~ NA_real_  # Default case (catches any unexpected values)
  ))
    

#Step 3: Institutional sector of economic activities
ra_id <- ra_id %>%
  mutate(ins_sector = case_when(
    employed== 1 & EMP10 %in% c(1,6)~ 1, #Public 
    employed== 1 & EMP10 %in% c(2,3,5)~ 2, #Private
    employed== 1 & EMP10 == 4 ~ 3, # Household domestic worker
    employed== 1 & (EMP10 %in% c(96,98) | is.na(EMP10)) ~ 4)) #Not classified

#Step 4: Employment benefits
#Social security contribution (proxy:pension fund)
ra_id <- ra_id %>%
  mutate(socsec = case_when(
  employed== 1 & EMP10_4 == 1  ~ 1, #Social security
  employed== 1 & EMP10_4 == 2  ~ 2, #No social security
  employed== 1 & (EMP10_4 == 98 | is.na(EMP10_4))  ~ 3)) #Not stated

#Paid leave
ra_id <- ra_id %>%
  mutate(painleave = case_when(
  employed== 1 & EMP10_5 == 1  ~ 1, #Paid leave
  employed== 1 & EMP10_5 == 2  ~ 2, #No paid leave
  employed== 1 & (EMP10_5 == 98 | is.na(EMP10_5))  ~ 3)) #Not stated

#Paid sick leave
ra_id <- ra_id %>%
  mutate(sickleave = case_when(
  employed== 1 & EMP10_6 == 1  ~ 1, #Sick leave
  employed== 1 & EMP10_6 == 2  ~ 2, #No sick leave
  employed== 1 & (EMP10_6 == 98 | is.na(EMP10_6))  ~ 3)) #Not stated

#Parental leave
ra_id <- ra_id %>%
  mutate(parentalleave = case_when(
  employed== 1 & EMP10_6a == 1  ~ 1, #Parental leave
  employed== 1 & EMP10_6a == 2  ~ 2, #No parental leave
  employed== 1 & (EMP10_6a == 98 | is.na(EMP10_6a))  ~ 3)) #Not stated

##Step 5: Characteristics of the enterprise
#Registration 
ra_id <- ra_id %>%
  mutate(registered = case_when(
  employed== 1 & EMP10_11 == 1 ~ 1, #Registered 
  employed== 1 & EMP10_11 == 2 ~ 2, #Not registered 
  employed== 1 & (EMP10_11 == 98 | is.na(EMP10_11)) ~ 3)) #Not stated 
#Bookkeeping 
ra_id <- ra_id %>%
  mutate(bookkeep = case_when(
  employed== 1 & EMP10_12 == 1 ~ 1, #Keep accounts for tax purposes - formal
  employed== 1 & EMP10_12 %in% c(2,3,4) ~ 2, #Keep accounts not for tax purposes; Personal use only; Does not keep accounts
  employed== 1 & (EMP10_12 == 98 | is.na(EMP10_12)) ~ 3)) #Not stated 

#Step 6: Informal/formal economy
#UNIT OF PRODUCTION: FORMAL/INFORMAL SECTOR 
ra_id <- ra_id %>%
  mutate(unit_prod = case_when(
    employed == 1 & ins_sector == 3 ~ 3, # Households
    employed == 1 & (ins_sector == 1 | registered == 1 | bookkeep == 1) ~ 2, # Employment in formal sector
    employed == 1 & !(ins_sector == 3 | ins_sector == 1 | registered == 1 | bookkeep == 1) ~ 1, # Informal sector
    TRUE ~ NA_real_ # Default case for missing values
  ))

#NATURE OF JOB: FORMAL/INFORMAL EMPLOYMENT 
ra_id <- ra_id %>%
  mutate(job_nature = case_when(
    (isce93  %in% c(2,3) & unit_prod !=2) | (isce93  %in% c(1, 5) & socsec !=1) | isce93 == 4 ~ 1, #Informal employment
    (isce93  %in% c(2,3) & unit_prod ==2) | (isce93  %in% c(1, 5) & (socsec ==1 | (socsec == 3 & painleave ==1 & sickleave ==1))) ~ 0)) #Formal employment 

```


```{r, eval=TRUE}
#Step 1: Determine who is employed based on various conditions 
ra_id <- ra_id %>%
  mutate(employed = case_when(
    EMP01 == 1  ~ 1,  # Paid employment - employees
    EMP02 == 1 & EMP06b == 3 ~ 1,  # Self-employment
    EMP03 == 1 & EMP06b == 3 ~ 1,  # Unpaid contributing family workers
    EMP04 == 1  ~ 1,  # Absent but still employed
    EMP05 == 1 & EMP06a == 3~ 1,  # Absent 
    EMP05 == 1 & EMP06b == 3~ 1,  # Absent self-employed
    EMP02 == 1 & EMP06b %in% c(1, 2) & EMP06c %in% c(1, 2) ~ 1,  # Farming/rearing/fishing for sale
    EMP03 == 1 & EMP06b %in% c(1, 2) & EMP06c %in% c(1, 2) ~ 1,  # ADDED: FAMILY farming/rearing/fishing for sale
    EMP05 %in% c(1, 2) & EMP06b %in% c(1, 2) & EMP06c %in% c(1, 2) ~ 1,  # Temporarily bsent but farming for sale
    EMP05 %in% c(1, 2) & EMP06a %in% c(1, 2) & EMP06b %in% c(1, 2) ~ 1,  # 
    TRUE ~ 0  # Not employed
  ))

#Step 2: Define unemployed: those not employed but actively looking for work and available to carry out activities
ra_id <- ra_id %>%
  mutate(unemployed = case_when(
    (employed == 0 & (EMP25a == 1 
                      #| #EMP25aa == 1 (excluded as not in SSD)
                      ) & EMP29 == 1) ~ 1,  # Actively looking for work
    TRUE ~ 0  # Not unemployed if not actively looking or employed
  ))

#Step 3: Define labour force: anyone employed or unemployed
ra_id <- ra_id %>%
  mutate(labour_force = case_when(
    (employed == 1 | unemployed == 1) ~ 1,  # Employed or actively seeking work
    TRUE ~ 0  # Not in labour force
  ))

ra_id <- ra_id %>%
  mutate(family_farming = case_when(
    EMP06b  %in% c(3,4) | EMP06c %in% c(3,4) ~ 1,  
    TRUE ~ 0
  ))

#Step 4: Outcome - Unemployment rate
ra_id <- ra_id %>%
  mutate(Unemployed = case_when(
    employed == 1 & labour_force == 1 ~ 0,  # If employed, outcome is 2 (not unemployed)
    unemployed == 1 & labour_force == 1 ~ 1,  # If unemployed, outcome is 1
    TRUE ~ NA_real_  # Catch-all for cases not in the labor force or with missing data
  ))

ra_id <- ra_id %>%
   mutate(Unemployed = labelled(Unemployed,
                                labels = c(
                                  "Unemployed" = 1,  # Label for Yes
                                  "Not unemployed" = 0   # Label for No
                                ),
                                label = "Proportion of people (working age) who are unemployed"))
```


**Food consumption score**
```{r}
roster_id <- roster_id %>%
  mutate(
    FCSStap   = Food11aa,  # Cereals/staples
    FCSPulse  = Food12a,   # Legumes/nuts
    FCSDairy  = Food13a,   # Milk/dairy
    FCSPr     = Food14a,   # Meat/fish/eggs
    FCSVeg    = Food15a,   # Vegetables
    FCSFruit  = Food16a,   # Fruits
    FCSFat    = Food17a,   # Oils/fats
    FCSSugar  = Food18a  ) %>%
  mutate(across(FCSStap:FCSSugar, as.numeric))

roster_id <- roster_id %>%
  mutate(
    FCS = FCSStap * 2 +      # Cereals weight: 2
          FCSPulse * 3 +     # Pulses weight: 3
          FCSDairy * 4 +     # Dairy weight: 4
          FCSPr * 4 +        # Protein weight: 4
          FCSVeg +       # Vegetables weight: 1
          FCSFruit +     # Fruits weight: 1
          FCSFat * 0.5 +       # Fats weight: 0.5
          FCSSugar * 0.5       # Sugar weight: 0.5
  )

roster_id <- roster_id %>%
  mutate(
    FCSCat21 = case_when(
      FCS <= 21 ~ "Poor",
      FCS <= 35 ~ "Borderline",
      FCS > 35 ~ "Acceptable",
      TRUE ~ NA_character_
    ),
    FCSCat28 = case_when(
      FCS <= 28 ~ "Poor",
      FCS <= 42 ~ "Borderline",
      FCS > 42 ~ "Acceptable", 
      TRUE ~ NA_character_
    )
  )

fcs_ci <- summary_ci(roster_id, "FCS")

roster_id <- roster_id |>
  mutate(
    FCSPoorBorderline28 = case_when(
      FCSCat28 %in% c("Poor", "Borderline") ~ 1,
      TRUE ~ 0
    )
  ) |>
  mutate(
    FCSPoorBorderline21 = case_when(
      FCSCat21 %in% c("Poor", "Borderline") ~ 1,
      TRUE ~ 0
    )
  )
```

```{r}
roster_id <- roster_id |> 
  mutate(
    `Social protection` = case_when(ScProtec01 == 1 ~ 1,
                                     TRUE ~ 0)
    ) |>
   mutate(`Social protection` = labelled(`Social protection`,
                                labels = c(
                                  "Social protection" = 1,
                                  "No social protection" = 0
                                ),
                                label = "Social protection"))
```


```{r}
summary_diff <- function(data,
                                         vars,
                                         ref_label = "Recognised refugees",
                                         criteria_var = "criteria_ref",
                                         country_var = "Country") {
  # Long format
  long <- data %>%
    select(all_of(c(country_var, criteria_var, vars))) %>%
    pivot_longer(cols = all_of(vars), names_to = "variable", values_to = "value")
  
  # Group summaries: by Country × criteria_ref × variable
  grp_sum <- long %>%
    group_by(.data[[country_var]], .data[[criteria_var]], variable) %>%
    summarise(
      n = sum(!is.na(value)),
      mean = mean(value, na.rm = TRUE),
      sd   = sd(value, na.rm = TRUE),
      se   = sd / sqrt(n),
      .groups = "drop"
    ) %>%
    rename(
      Country = !!country_var,
      criteria_ref = !!criteria_var
    )
  
  # Reference group per Country × variable
  ref_sum <- grp_sum %>%
    filter(criteria_ref == ref_label) %>%
    rename(mean_ref = mean, sd_ref = sd, se_ref = se, n_ref = n) %>%
    select(Country, variable, mean_ref, sd_ref, se_ref, n_ref)
  
  # Other groups
  oth_sum <- grp_sum %>%
    filter(criteria_ref != ref_label) %>%
    rename(mean_g = mean, sd_g = sd, se_g = se, n_g = n)
  
  # Comparisons vs reference within Country × variable
  cmp <- oth_sum %>%
    left_join(ref_sum, by = c("Country", "variable")) %>%
    mutate(
      diff_mean = mean_g - mean_ref,
      se_diff   = sqrt(se_g^2 + se_ref^2),
      z_diff    = diff_mean / se_diff,
      ci_low    = diff_mean - 1.96 * se_diff,
      ci_high   = diff_mean + 1.96 * se_diff,
      # Standardized Mean Difference (Cohen's d)
      sd_pooled = sqrt(((pmax(n_g - 1, 1)) * sd_g^2 + (pmax(n_ref - 1, 1)) * sd_ref^2) / pmax(n_g + n_ref - 2, 1)),
      smd       = diff_mean / sd_pooled,
      # SD differences
      diff_sd   = sd_g - sd_ref,
      sd_ratio  = sd_g / sd_ref
    ) %>%
    select(Country, variable, criteria_ref,
           n_g, mean_g, sd_g,
           n_ref, mean_ref, sd_ref,
           diff_mean, z_diff, ci_low, ci_high, smd, diff_sd, sd_ratio)
  
  list(
    group_summary = grp_sum,
    comparisons = cmp
  )
}



vars_to_check <- c("FCSPoorBorderline28", "Land documents", "Land tenure", "Social protection")

out <- summary_diff(
  data = roster_id,
  vars = vars_to_check,
  ref_label = "Refugee based on criteria",  # your chosen reference label
  criteria_var = "criteria_ref",
  country_var = "Country"
)

vars_to_check_ra <- c("Financial account", "Unemployed")

out_ra <- summary_diff(
  data = ra_id,
  vars = vars_to_check_ra,
  ref_label = "Refugee based on criteria",  # your chosen reference label
  criteria_var = "criteria_ref",
  country_var = "Country"
)

out$comparisons <- bind_rows(out$comparisons, out_ra$comparisons)
# Browse outputs
dplyr::arrange(out$group_summary, Country, variable, criteria_ref) %>% head()
dplyr::arrange(out$comparisons, Country, variable, criteria_ref) %>% head()


cmp <- out$comparisons %>%
  mutate(
    panel = Country,
    p_value = 2 * pnorm(abs(z_diff), lower.tail = FALSE),
    sig = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      p_value < 0.1   ~ ".",
      TRUE            ~ ""
    )
  )

# Helpers to reorder within facets
reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  x <- as.character(x); within <- as.character(within)
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun, ...)
}
scale_y_reordered <- function(sep = "___") {
  scale_y_discrete(labels = function(x) gsub(paste0(sep, ".*$"), "", x))
}
```

```{r}
ggplot(
  cmp,
  aes(
    x = z_diff,
    y = Country,
    color = sig
  )
) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high), height = 0, size = 0.6, alpha = 0.85) +
  geom_point(size = 2) +
  facet_grid(~ variable, scales = "free_y", space = "free_y") +
  scale_y_reordered() +
  labs(
    x = NULL,
    y = NULL,
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "bottom"
  )
```

# Roster position
We test whether the position in the roster has an effect on the likelihood of completing the roster. T
```{r, fig.cap = "Share of refugees with a valid document completing the identification questions, by member position and relation to household head"}

top5_relationships <- roster_id %>%
  filter(!HH_03 %in% c("The household head himself/herself", "Household head", "Wife/Husband/Partner")) %>%
  count(HH_03, sort = TRUE) %>%
  slice_head(n = 2)


plot_data <- roster_id  %>%
  filter(HH_03 %in% top5_relationships$HH_03) %>%
  mutate(HH_03 = labelled::to_factor(HH_03)) %>%
  filter(as.numeric(mPosition) %in% c(2:7)) %>%
  group_by(Country, mPosition, HH_03) %>%
  summarise(
    n = n(),
    proportion = mean(criteria_ref_num == 1)
  ) %>%
  mutate(
    se = sqrt(proportion * (1 - proportion) / n),
    ci_lower = proportion - 1.96 * se,
    ci_upper = proportion + 1.96 * se,
    proportion_pct = proportion * 100,
    ci_lower_pct = ci_lower * 100,
    ci_upper_pct = ci_upper * 100
  ) 

# Plot

ggplot(plot_data, aes(x = as.integer(mPosition), y = proportion_pct, color = HH_03, group = HH_03)) +
  geom_line() +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = ci_lower_pct, ymax = ci_upper_pct), width = 0.2, alpha = .2) +
  scale_y_continuous(labels = percent_format(scale = 1), limits = c(0, 100)) +
  geom_hline(yintercept = 0) +
  labs(
    x = "Roster position",
    y = ""
  ) +
  theme_unhcr( grid = "Y") +
  facet_grid(HH_03 ~ Country)



```

